---
title: "class1-b"
author: "Sujatha Jagannathan"
date: "February 24, 2020"
output: ioslides_presentation
incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Good coding practices

>   "Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread."
>   
>   --- Hadley Wickham

## File Names

File names should be meaningful and end in `.R`, `.Rmd`, etc. Avoid using special characters in file names - stick with numbers, letters, `-`, and `_`. *Never* include spaces in file names! 

    ###### Good  
    fit_models.R  
    utility_functions.Rmd  

    ###### Bad  
    fit models.R  
    foo.r  
    stuff.r  

## File Names

If files should be run in a particular order, prefix them with numbers. If it 
seems likely you'll have more than 10 files, left pad with zero. It looks nice (constant width) and sorts nicely. 

    00_download.R  
    01_explore.R  
    ...  
    09_model.R  
    10_visualize.R  
    
## File Names

> Avoid capitalizing when not necessary. 

> If you want to include dates in your file name, use the ISO 8601 standard: `YYYY-MM-DD`

> Use delimiters intentionally! (helps you to recover metadata easily from file names)

    "_" delimits fields
    "-" delimits words
    
    Example: 2019-02-15_class1_data-wrangling.Rmd

## File Names

Avoid hard coding file names and instead use relative paths. Use `getwd()` to figure out what your working directory is. `~` represents the current working directory. 

    ###### Good
    "~/class1/code/test.R"
    
    ###### Bad
    "/Users/sjaganna/Desktop/CU-onedrive/08-teaching/molb7910/class1/data.csv"

## Organisation

> Try to give a file a concise name that evokes its contents

One way to organize your files is by grouping them into `data`, `code`, `plots`, etc.   

## Organisation

For **this class** PLEASE use the following structure:  

      class1
         - data
          + input
            - counts.csv
            - metadata.txt
          + output
            - counts_mod.csv
        - code
          + explore.Rmd
          + visualize.Rmd
        - images
          + plot1
          + plot2

## Internal structure 

Use commented lines of `-` and `=` to break up your code chunk into easily readable 
segments. Or better yet, make each "action" it's own chunk and give it a name. 

```{r, eval = FALSE}
# Load data ---------------------------

# Plot data ---------------------------
```

## Packages

If your script uses add-on packages, load them all at once at the very 
beginning of the file. This is more transparent than sprinkling `library()` 
calls throughout your code or having hidden dependencies that are loaded in a 
startup file, such as `.Rprofile`.

## Acknowledgements
Source: https://style.tidyverse.org/
