---
title: "Homework Class 7"
author: "Neelanjan Mukherjee"
date: "03/08/2020"
output: html_document
---
<br/>

The goal of this homework is to perform analysis of genomic data. 

Please execute the following tasks, knit the resulting .Rmd (Lastname_class7-homework.Rmd) and submit the .html output as your assignment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE)
```

<br/>

#### Load packages 

```{r, include = TRUE}
#load packages here

```

<br/>

## Answer the following questions
### Examine HuR binding sequence specificity
#### What sequences does HuR preferentially bind to in introns vs intronic background?

Perform the same analysis on introns that was performed in 3' UTRs in class.


```{r intron seqs}

# get intron sequences
# get  sequences of hur binding sites in introns
# count 6 mer frequencies in both
# plot the difference in 6mer freq between intronic hur bound seqs and intron seqs
# name top 3 enriched 6mers


```



#### Session info
Finally, we will __always__ end our Rmarkdown documents with the sessing information:

```{r}
# session info
```
