---
title: "class3-workbook"
author: "Sujatha Jagannathan"
date: "2/27/2020"
output: html_document
---

## Load packages: tidyverse, ggthemes, scales, cowplot, viridis

```{r setup}
library(tidyverse) # data-wrangling + ggplot2
library(ggthemes) # pretty plotting themes
library(scales) # nice annotating of x/y axes in plots
library(cowplot) # to make panels of plots
library(viridis) # nice colors!
```

## No need to `setwd()`. But check with `getwd()`
```{r}
getwd()
```

## Import data
```{r}
data_gene_level <- read_csv("data/input/data_gene_level.csv")
rp_genes <- read_csv("data/input/rp_genes.csv")
target_genes <- read_csv("data/input/target_genes.csv")
```

## Add DUX4 target status using `ifelse()`
```{r}
# if-else clause within mutate
data <- data_gene_level %>%
          mutate(target_status = 
                ifelse(hgnc_symbol %in% target_genes$hgnc_symbol, 
                       "target", "non-target"))
```

## Example: making a plot step-by-step

```{r}
# initialize with data
ggplot(data)

# specify x and y axes
ggplot(data, aes(x = time, y = count))
# specify geometry
ggplot(data, aes(x = time, y = count)) + 
  geom_boxplot()

```

## transform data within ggplot
```{r}
ggplot(data, aes(x = time, y = log10(count))) + 
  geom_boxplot()

# add pseudocount of 10
ggplot(data, aes(x = time, y = log10(count+10))) + 
  geom_boxplot()
```

## Preserving the order of samples
```{r}
data$time <- factor(data$time, 
                    levels = c("0h", "4h", "8h", "14h"))

# let's try the plot again...
ggplot(data, aes(x = time, y = log10(count+10))) + 
  geom_boxplot()
```

## changing the theme
```{r}
p <- ggplot(data, aes(x = time, y = log10(count+10))) + 
  geom_boxplot()

# theme_few
p + theme_few()
# theme_minimal
p + theme_minimal()
# theme_wsj
p + theme_wsj()

```

## Adding plot title
```{r}
p + theme_few() + ggtitle("DUX4 time course")
```

## Modifying axis labels
```{r}
p + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab("RNA-seq read counts") + 
  xlab("Time of DUX4 expression")
```

## Let's explore other geometries - jitter plot

```{r}
p <- ggplot(data, 
            aes(x = time, y = log10(count+10))) + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab("RNA-seq read counts") + 
  xlab("Time of DUX4 expression")

# geom_point

p + geom_point()

# let's jitter the points
p + geom_point(position = "jitter")
# let's make the points more transparent
p + geom_point(position = "jitter", 
               alpha = 0.01, 
               color = "red")
```

## Let's explore other geometries - violin plot

```{r}
# geom_violin
p + geom_violin() + geom_point()
# let's jitter the points

```

## let's make a histogram
```{r}
# geom_histogram
ggplot(data, aes(log10(count+ 10))) + geom_histogram(fill = "#ffeeee")

# why can't I do this?
hist(log10(data$count+10))

# you can, while exploring data. ggplot is better to make figures since it gives you easy control of the aesthetics. 
```

## let's go back to box plot
```{r}
p <- ggplot(data, aes(x = time, y = log10(count+10))) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  

p + geom_boxplot()
```

## now let's play with color
```{r}
# changing the outline and fill color
p + geom_boxplot(color = "black", fill = "red", cex = 2)
```

## what if you want to color by another variable?
```{r}
# add a `fill` to your aesthetic
p <- ggplot(data, aes(x = time, y = log10(count+10), fill = time)) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression") 

p + geom_boxplot()
```

## plotting by different groups
```{r}
# fill by target_status
p <- ggplot(data, aes(x = time, y = log10(count+10), fill = target_status)) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression") 

# change up x and fill
p <- ggplot(data, aes(x = target_status, y = log10(count+10), fill = time)) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression") 

p + geom_boxplot()
```

## default colors are boring
```{r}
# use viridis
p_new <-
p + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, begin = .2, end = .8) 

p_new


```

## What else can ggplot do?

A LOT more! Before we go into that, let's look under the hood of a ggplot object. 

```{r}
#let's save our plot as an object
plot <-
p + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, begin = .2, end = .8) 

```

## what is facet? 

```{r}
# `facet_wrap` time
plot + facet_wrap(~time)
# `facet_wrap` target_status
plot + facet_wrap(~target_status)
# `facet_wrap` replicate
plot + facet_wrap(~replicate)
# `facet_wrap` multiple variables
plot + facet_wrap(time~replicate)
# `facet_grid` 
plot + facet_grid(time~replicate)

# let's fix x-axis

```

## How to combine multiple plots into a figure?
```{r}
# plot1
plot1 <- plot + facet_wrap(~time)

# plot2 
plot2 <- plot + facet_wrap(~replicate)

# Now let's make a figure!
plot_final <- plot_grid(plot1, plot2, 
                        labels = c("A", "B"),
                        nrow = 2)
```

## Saving plots
```{r}
save_plot(filename = "images/dux4_expression_time.png", 
          plot = plot1, 
          base_width = 6, 
          base_height = 4)

save_plot(filename = "images/dux4_expression_replicate.png", 
          plot = plot2, 
          base_width = 6, 
          base_height = 4)

save_plot(filename = "images/final_figure.png", 
          plot = plot_final, 
          base_width = 6, 
          base_height = 4)
```

## session information
```{r}
sessionInfo()
```
