---
title: "class3-presentation"
author: "Sujatha Jagannathan"
date: "February 28, 2019"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)  # turn off scientific notation like 1e+06
```

## Learning goals

- Understand the syntax of ggplot2
- Get comfortable with plotting, layering features to create complex plots, making "publication ready" figures including panels and labels
- Use in-class time to work through issues and answer questions

## Load packages

```{r setup,echo=FALSE, message=FALSE}

# We need to load all the libraries we think are needed for this exercise.
library(tidyverse) # data-wrangling + ggplot2
library(ggthemes) # pretty plotting themes
library(scales) # nice annotating of x/y axes in plots
library(cowplot) # to make panels of plots
library(viridis) # nice colors!
```

## Set working directory
```{}
No need to do this today because we are going to try something new. Please open the `class3-workbook.Rmd` from the `3-class3` folder you just downloaded. The default behavior of Rstudio is to consider the folder that contains a .Rmd document as the working directory. We are going to take advantage of this and see if this fixes our knitting and file path issues. All relative paths we will use today will be with respect to the `3-class3` folder that contains the .Rmd file. 
```

## Import data
```{r}
data_gene_level <- read_csv("data/input/data_gene_level.csv")
rp_genes <- read_csv("data/input/rp_genes.csv")
target_genes <- read_csv("data/input/target_genes.csv")
```

## Add DUX4 target status
```{r}
# if-else clause within mutate
data <- data_gene_level %>%
          mutate(target_status = 
                ifelse(hgnc_symbol %in% target_genes$hgnc_symbol, "target", "non-target"))
```

## 1. Plotting data with ggplot2 

A popular approach to plotting in R is [ggplot2](https://ggplot2.tidyverse.org/). While graphs produced using ggplot2 defaults are (imo) not quite publication qualitym with a few tweaks you can pretty much get there! I find [STHDA](http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization) to be a great resource for `ggplot2` and many other `R` coding questions.

## The basic syntax of `ggplot()`  

`ggplot()`: build plots piece by piece
The concept of ggplot divides a plot into three different fundamental parts:

plot = __data__ + __aesthetics__ + __geometry__.  
<br/>
__data__: a data frame
<br/>
__aesthetics__: specify x and y variables, and other features - color, size, shape, etc.
<br/>
__geometry__: specify type of plots  - histogram, boxplot, line, density, dotplot, bar, etc. 
<br/>


## Example: making a plot step-by-step
We will make a variety of box plots now. Please see this [webpage](http://www.sthda.com/english/wiki/ggplot2-stripchart-jitter-quick-start-guide-r-software-and-data-visualization) for more information.

```{r}
# initialize with data
ggplot(data) # specify which dataframe to use - no plot yet

# specify x and y axes
ggplot(data, aes(x = time, y = count)) # `data` is 'data', `x` is 'time', `y` is 'count'

# specify geometry
ggplot(data, aes(x = time, y = count)) +  
  geom_boxplot() # `geom` is 'geom_boxplot()'
```

## transform data within ggplot
```{r}
# first method
ggplot(data, aes(x = time, y = log10(count))) +  
  geom_boxplot() 

# second method
ggplot(data, aes(x = time, y = count)) +  
  geom_boxplot() +
  scale_y_log10() # make y-axis log10

# add pseudocount of 10
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() # make y-axis log10
```

## Preserving the order of samples
```{r}
data$time <- factor(data$time, levels = c("0h", "4h", "8h", "14h"))

# let's try the plot again...
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() # make y-axis log10
```

## changing the theme
```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() + 
  theme_few() # fewer lines

ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() + 
  theme_minimal() # Minimal

ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() + 
  theme_fivethirtyeight() # Nate Silver + Co.
```

## Adding plot title
```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() +
  theme_few() +
  ggtitle("DUX4 time course") # title
```

## Modifying axis labels
```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + # my new y-axis name
  xlab(label = "Hours of DUX4 expression")  # my new x-axis name

```

## Let's explore other geometries - jitter plot

```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_point() +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  

# let's jitter the points
ggplot(data, aes(x = time, y = count+10)) +  
  geom_point(position = "jitter") +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression") 

# let's make the points more transparent
ggplot(data, aes(x = time, y = count+10)) +  
  geom_point(position = "jitter", alpha = 0.01) +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
```

## Let's explore other geometries - violin plot

```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_violin() +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  

# let's jitter the points
ggplot(data, aes(x = time, y = count+10)) +  
  geom_violin() +
  geom_point(position = "jitter", alpha = 0.01) +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
```

## let's make a histogram
```{r}
ggplot(data, aes(count+10)) +  # only one variable specified
  geom_histogram() + # histogram
  scale_x_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "Frequency") + 
  xlab(label = "log10(count+10")  

# why can't I do this?
hist(log10(data$count+10))

# you can, while exploring data. ggplot is better to make figures since it gives you easy control of the aesthetics. 
```

## let's go back to box plot
```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot() +
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
```

## now let's play with color
```{r}
ggplot(data, aes(x = time, y = count+10)) +  
  geom_boxplot(color = "black", fill = "red") + # changing the outline and fill color
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
```

## what if you want to color by another variable?
```{r}
ggplot(data, aes(x = time, y = count+10, fill = replicate)) +  
  geom_boxplot() + 
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")
```

## plotting by different groups
```{r}
ggplot(data, aes(x = replicate, y = count+10, fill = time)) +  
  geom_boxplot() + 
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")

ggplot(data, aes(x = target_status, y = count+10, fill = time)) +  
  geom_boxplot() + 
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")
```

## default colors are boring
```{r}
ggplot(data, aes(x = target_status, y = count+10, fill = time)) +  
  geom_boxplot() + 
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression") + 
  scale_fill_viridis(discrete = TRUE) 
```

## What else can ggplot do?

A LOT more! Before we go into that, let's look under the hood of a ggplot object. 

```{r round two jitter}
#instead of plotting directly, let's save our plot as an object
plot <- ggplot(data, aes(x = target_status, y = count+10, fill = time)) +  
  geom_boxplot() + 
  scale_y_log10() + 
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression") + 
  scale_fill_viridis(discrete = TRUE) 

plot # prints plot
names(plot) # what is in the object?
plot$data # i can get the data back
plot$mapping # my aesthetics
plot$labels # i see my labels
plot$facet
```

## what is facet? 
let's say I want to make a separate `p_target`  plot for each gene

```{r}
plot + facet_wrap(~time) # `facet_wrap` time
plot + facet_wrap(~target_status) # `facet_wrap` target_status
plot + facet_wrap(~replicate) # `facet_wrap` replicate

# let's fix x-axis
plot + facet_wrap(~replicate) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree
```

## How to combine multiple plots into a figure?
```{r}
plot1 <- plot + facet_wrap(~time)
plot2 <- plot + facet_wrap(~replicate) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # angle 60 degree

# Now we make a figure!
plot_final <- plot_grid(plot1, plot2, labels = c("A","B"), nrow = 2)

plot_final
```

## Saving plots
```{r}
save_plot(filename = "images/dux4_time_course_target.png", plot = plot1, base_width = 6, base_height = 4)

save_plot(filename = "images/dux4_time_course_replicate.png", plot = plot2, base_width = 6, base_height = 4)

save_plot(filename = "images/figure2.png", plot = plot_final, base_width = 8, base_height = 8)
```

## session information
```{r}
sessionInfo()
```
