---
title: "class3-presentation"
author: "Sujatha Jagannathan"
date: "February 28, 2019"
output: ioslides_presentation

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)  # turn off scientific notation like 1e+06
```

## Learning goals

- Understand the syntax of ggplot2
- Get comfortable with plotting, layering features to create complex plots, making "publication ready" figures including panels and labels
- Use in-class time to work through issues and answer questions

## Load packages

```{}
# We need to load all the libraries we think are needed for this exercise.
library(tidyverse) # data-wrangling + ggplot2
library(ggthemes) # pretty plotting themes
library(scales) # nice annotating of x/y axes in plots
library(cowplot) # to make panels of plots
library(viridis) # nice colors!
```

## Set working directory

- No need to do this today because we are going to try something new. 
- Please open the `class3-workbook.Rmd` from the `3-class3` folder you just downloaded. 
- The default behavior of Rstudio is to consider the folder that contains a .Rmd document as the working directory. 
- We are going to take advantage of this and see if this fixes our knitting and file path issues. 
- All relative paths we will use today will be with respect to the `3-class3` folder that contains the .Rmd file. 

## Import data
```{}
data_gene_level <- read_csv("data/input/data_gene_level.csv")
rp_genes <- read_csv("data/input/rp_genes.csv")
target_genes <- read_csv("data/input/target_genes.csv")
```

## Add DUX4 target status
```{}
# if-else clause within mutate
data <- data_gene_level %>%
          mutate(target_status = 
                ifelse(hgnc_symbol %in% target_genes$hgnc_symbol, 
                "target", "non-target"))
```

## 1. Plotting data with ggplot2 

A popular approach to plotting in R is [ggplot2](https://ggplot2.tidyverse.org/). While graphs produced using ggplot2 defaults are not quite publication quality, with a few tweaks you can pretty much get there! 

I find [STHDA](http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization) to be a great resource for `ggplot2` and many other `R` coding questions.

## The basic syntax of `ggplot()`  

`ggplot()`: build plots piece by piece
The concept of ggplot divides a plot into three different fundamental parts:

plot = __data__ + __aesthetics__ + __geometry__.  
<br/>
__data__: a data frame
<br/>
__aesthetics__: specify x and y variables, and other features - color, size, shape, etc.
<br/>
__geometry__: specify type of plots  - histogram, boxplot, line, density, dotplot, bar, etc. 
<br/>


## Example: making a plot step-by-step

```{}
# initialize with data
ggplot(data) # specify which dataframe to use - no plot yet

# specify x and y axes
ggplot(data, aes(x = time, y = count)) # `data` is 'data', `x` is 'time', `y` is 'count'

# specify geometry
ggplot(data, aes(x = time, y = count)) +  
  geom_boxplot() # `geom` is 'geom_boxplot()'
```

## transform data within ggplot
```{}
ggplot(data, aes(x = time, y = log10(count))) +  
  geom_boxplot() 

# add pseudocount of 10
ggplot(data, aes(x = time, y = log10(count+10))) +  
  geom_boxplot() 
```

## Preserving the order of samples
```{}
data$time <- factor(data$time, levels = c("0h", "4h", "8h", "14h"))

# let's try the plot again...
ggplot(data, aes(x = time, y = log10(count+10))) +  
  geom_boxplot() 
```

## changing the theme

There are a lot of themes to choose from within `ggthemes`. You can take a look at this [quick guide](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
```{}
p <- ggplot(data, aes(x = time, y = log10(count+10))) +  
  geom_boxplot()

p + theme_few() # fewer lines

p + theme_minimal() # Minimal

p + theme_wsj() # Wall Street Journal
```

## You can also customize pre-existing themes

```{}
mytheme <- theme_minimal(base_size = 15) + 
                theme(aspect.ratio = 1, 
                      panel.background = 
                          element_rect(
                              colour = "black", 
                              size=1))

p + mytheme
```

## Adding plot title
```{}
p + theme_few() +
  ggtitle("DUX4 time course") # title
```

## Modifying axis labels
```{}
p + theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + # my new y-axis name
  xlab(label = "Hours of DUX4 expression")  # my new x-axis name
```

## Let's explore other geometries - jitter plot

```{}
p <- ggplot(data, aes(x = time, y = log10(count+10))) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
  
p + geom_point()
```

## let's jitter the points
```{}
p + geom_point(position = "jitter")
```

## let's make the points more transparent
```{}
p + geom_point(position = "jitter", alpha = 0.01)
```

## Let's explore other geometries - violin plot

```{}
p + geom_violin() 
```

## let's jitter the points
```{}
p + geom_violin() +
  geom_point(position = "jitter", alpha = 0.01) 
```

## let's make a histogram
```{}
p <- ggplot(data, aes(log10(count+10))) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  

p + geom_histogram() # histogram
```

## why can't I just do this?
`hist(log10(data$count+10))`

you can, while exploring data. ggplot is better to make figures since it gives you easy control of the aesthetics. 


## let's go back to box plot
```{}
p <- ggplot(data, aes(x = time, y = log10(count+10))) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
  
p + geom_boxplot()
```

## now let's play with color
```{}
# changing the outline and fill color
p + geom_boxplot(color = "black", fill = "red")  

```

## what if you want to color by another variable?
```{}
p <- ggplot(data, aes(
                  x = time, 
                  y = log10(count+10), 
                  fill = replicate)) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
  
p + geom_boxplot()
```

## plotting by different groups
```{}
p <- ggplot(data, aes(
                  x = time, 
                  y = log10(count+10), 
                  fill = target_status)) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
  
p + geom_boxplot()
```

## plotting by different groups
```{}
p <- ggplot(data, aes(
                  x = target_status, 
                  y = log10(count+10), 
                  fill = time)) +  
  theme_few() + 
  ggtitle("DUX4 time course") + 
  ylab(label = "RNA-seq read counts") + 
  xlab(label = "Hours of DUX4 expression")  
  
p + geom_boxplot()
```

## default colors are boring
```{}
p + geom_boxplot() + 
  scale_fill_viridis(discrete = TRUE) 
```

## What else can ggplot do?

A LOT more! Before we go into that, let's look under the hood of a ggplot object. 

```{}
plot <- p + geom_boxplot() + 
  scale_fill_viridis(discrete = TRUE) 
```

## 
```{}
plot # prints plot
names(plot) # what is in the object?
plot$data # i can get the data back
plot$mapping # my aesthetics
plot$labels # i see my labels
plot$facet
```

## what is facet? 

```{}
plot + facet_wrap(~time) # `facet_wrap` time
plot + facet_wrap(~target_status) # `facet_wrap` target_status
plot + facet_wrap(~replicate) # `facet_wrap` replicate
plot + facet_wrap(time ~ replicate) # `facet_wrap` multiple
plot + facet_grid(time ~ replicate) # `facet_grid` 

# let's fix x-axis
plot + facet_grid(time ~ replicate) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) 
  # angle 60 degree
```

## How to combine multiple plots into a figure?
```{}
plot1 <- plot + facet_wrap(~time)
plot2 <- plot + facet_wrap(~replicate) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) 

# Now let's make a figure!
plot_final <- plot_grid(plot1, plot2, labels = c("A","B"), nrow = 2)

plot_final
```
More information on using plot_grid (from package `cowplot`) is [here](https://wilkelab.org/cowplot/articles/plot_grid.html)

## Saving plots
```{}
save_plot(filename = "images/dux4_time_course_target.png", 
          plot = plot1, 
          base_width = 6, 
          base_height = 4)

save_plot(filename = "images/dux4_time_course_replicate.png", 
          plot = plot2, 
          base_width = 6, 
          base_height = 4)

save_plot(filename = "images/figure2.png", 
          plot = plot_final, 
          base_width = 8, 
          base_height = 8)
```

## session information
```{r}
sessionInfo()
```
